# Data sources

We obtain three main datasets from NYC Open Data. Open Data is free public data published by New York City agencies and other partners, which collects all kinds of data in NYC.

In our project, we mainly consider about three types of datasets: COVID-19 datasets, which contain records of cases, deaths and vaccination populations; Crime datasets, which contain records of shooting, arrest and hate crime data in NYC; Air Quality dataset, which contains air quality data.

We will try to gain some insights from the above three different perspectives. We want to estimate the overall living condition across neighborhoods in New York City. After analyzing all these three perspectives, hopefully we may offer some advice on choosing the ideal neighborhood for people to live.

## COVID-19 Dataset

### Vaccinated Population Data.

These data show the percent of NYC residents who received at least one dose of COVID-19 vaccine and the percent who are fully vaccinated, by borough of residence and demographic group. The percentages reflect the number of people vaccinated within that specific demographic group. Vaccination rates are reported by standard U.S Census age categories. The population estimate available for youth eligible for vaccination is among ages 5 to 17.

Vaccination is one of the most effective ways to prevent COVID infection. With higher vaccination rate, community immunity will be more likely to be achieved. Thus if you want to find a place in new york, the vaccination rate can be taken into considerations. These indicators provide a perspective across time and NYC geographies to better characterize  health in NYC.


Vaccinated Population data contains these following key variables: 

```{r, echo  = FALSE}
Variable.name = c('BOROUGH','AGE_GROUP','RACE_ETHNICITY','PERC_FULLY','PERC_1PLUS')
Description = c('Borough of residence', 'Race/ethnicity group','Age group in years','Estimated percentage of people fully vaccinated', 'Estimated percentage of people vaccinated')
d = data.frame(Variable.name,Description)
knitr::kable(d, caption = "Key Variables in Air Quality Dataset",
             row.names = F,font_size = 10)
```


This dataset can be found on: https://github.com/nychealth/covid-vaccine-data/tree/main/people#coverage-summary-allagescsv


## Crime Dataset

### Shooting Incident Data collected by NYPD.
Such dataset can be found on: https://data.cityofnewyork.us/Public-Safety/NYPD-Shooting-Incident-Data-Historic-/833y-fsy8
https://data.cityofnewyork.us/Public-Safety/NYPD-Shooting-Incident-Data-Year-To-Date-/5ucz-vwe8

Key variables: `PERP_AGE_GROUP`, `PERP_SEX`, `VIC_AGE_GROUP`, `VIC_SEX`, and `STATISTICAL_MURDER_FLAG` across `BORO` at `OCCUR_TIME`.

### Arrest Data collected by NYPD.
Such dataset can be found on: https://data.cityofnewyork.us/Public-Safety/NYPD-Arrests-Data-Historic-/8h9b-rp9u
https://data.cityofnewyork.us/Public-Safety/NYPD-Arrest-Data-Year-to-Date-/uip8-fykc

Key variables: `LAW_CODE`, `AGE_GROUP`, `PERP_SEX`, and `PERP_RACE` across `ARREST_BORO` at `Record Create Date`.

### Hate Crimes Data collected by NYPD.
Such dataset can be found on: https://data.cityofnewyork.us/Public-Safety/NYPD-Hate-Crimes/bqiq-cu78

Key variables: `Law Code Category Description`, `Bias Motive Description`, and `Offense Description` across `County` at `ARREST_DATE`.

## Air Quality Dataset

### New York City Air Quality Surveillance Data:

This dataset contains information on New York City air quality surveillance data.

Air pollution is one of the most important environmental threats to urban populations and while all people are exposed, pollutant emissions, levels of exposure, and population vulnerability vary across neighborhoods. Exposures to common air pollutants have been linked to respiratory and cardiovascular diseases, cancers, and premature deaths. These indicators provide a perspective across time and NYC geographies to better characterize air quality and health in NYC.

Air quality is an important indicator of living wellness. Analyzing air quality data will help us to find a health place for people to live.

Air quality data contains these following key variables: 

```{r, echo  = FALSE}
Variable.name = c('Name', 'Measure', 'Geo Join ID', 'Start_Date', 'Data Value')
Description = c('Name of the indicator', 'How the indicator is measured','Identifier of the neighborhood geographic area','Date value for the start of the time_period', 'The actual data value for this indicator')
d = data.frame(Variable.name,Description)
knitr::kable(d, caption = "Key Variables in Air Quality Dataset",
             row.names = F,font_size = 10)
```
This dataset can be found on: https://data.cityofnewyork.us/Environment/Air-Quality/c3uy-2p5r
